Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-06 app-datastruct.tex
A few more edites to section 2. Also unified "time step" -> "timestep" throughout the whole manuscript.

\subsection{Creation and Destruction: \tt{RebuildHierarchy}}
Creation and destruction of grids happens in the routine
{\tt RebuildHierarchy}.  
At the end of every timestep, the Hierarchy is rebuilt.
Rebuilding starts at the coarsest level at the point in time (note
that coarser levels may be further ahead in their integration, due to
their larger time step and the W-cycle)  Beginning with this coarsest level, each
grid on each level is examined for refinement.  Cells are first flagged
based on a variety of criterion, described in section
\ref{sec.flagging}.  The set of flagged cells, stored in the \grid\ memeber
array {\tt FlaggingField}, is then devided into grids as described
in section \ref{app.subgrid_creation}.  Data in these subgrids
is then created by interpolation from the parent using \red{what kind
  of interpolation?}  After this New Subgrid is filled with
interpolated data, the set of Old Subgrids is examined for overlapping
data.  All Old Cells that correspond to New Cells are coppied to the
New Grid.

>>>>

\subsection{Creation and Destruction: \tt{RebuildHierarchy}}
Creation and destruction of grids happens in the routine
{\tt RebuildHierarchy}.  
At the end of every timestep, the Hierarchy is rebuilt.
Rebuilding starts at the coarsest level at the point in time (note
that coarser levels may be further ahead in their integration, due to
their larger timestep and the W-cycle)  Beginning with this coarsest level, each
grid on each level is examined for refinement.  Cells are first flagged
based on a variety of criterion, described in section
\ref{sec.flagging}.  The set of flagged cells, stored in the \grid\ memeber
array {\tt FlaggingField}, is then devided into grids as described
in section \ref{app.subgrid_creation}.  Data in these subgrids
is then created by interpolation from the parent using \red{what kind
  of interpolation?}  After this New Subgrid is filled with
interpolated data, the set of Old Subgrids is examined for overlapping
data.  All Old Cells that correspond to New Cells are coppied to the
New Grid.

### 2013-06-06 numerical-overview.tex
A few more edites to section 2. Also unified "time step" -> "timestep" throughout the whole manuscript.

Algorithmically, the star formation events in Kravtsov's code
are assumed to occur once every global time step 
$\Delta t_0 \leq 10^7$.  In cells where star formation is
determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
collisionless ``star particle'' is assumed to form, with a 
mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
volume of the mesh cell.  These star particles are
formed instantaneously, with all $m_*$ of gas being
taken out of the cell and immediately deposited into the
star particle.  This particle is then given the velocity
of the gas in the cell, and thereafter treated as a 
collisionless particle.  The \enzo\ implementation
of this algorithm is similar, except that instead of 
forming stars only at the root grid time step, we allow 
stars to form at the time step of the highest level of
resolution at any particular point in space.  As can be seen
from the equation for $m_*$ above, this can result in very 
small stellar masses.  To avoid memory and processor time
issues related to having very large numbers of star particles
we impose a threshold mass $M_{*,min}$ such that a star
particle only forms if $m_* \geq M_{*,min}$.  An
appropriate choice of $M_{*,min}$ does not significantly 
change the star overall star formation history of a 
simulation, though it may delay the onset of star formation 
in a given cell relative to a simulation without a particle mass
threshold.

@@

From the point of view of the hydro solver, each grid is solved as an
independent CFD problem, with Dirichlet boundary conditions stored in
the ghost zones.  At the end of each time step, cells on a given level
fill ghost zones by interpolating from the parent and copying from
neighbors on the same level.  Grids then correct the flux at the interface boundary between
fine and coarse zones, and finally project their real zone data to the
parent grid.

@@

\subsection{Time Stepping}\label{sec.timestepping}
\dcc{Made this its own section (not really part of the AMR, it's more
  general than that.)  Added stuff on harmonic average for 3d. See
  note below on acceleration timestep criterion being crap. Moved
  w-cycle to the previous section}
In \enzo, resolution of the equations being solved is adaptive in time 
as well as in space.  The timestep in \enzo\ is satisfied on a level-by-level
 basis by finding the largest timestep such that multiple criteria are
satisfied on each level.  The timestep criteria used by \enzo\ are 
(showing the one-dimensional case for clarity):

@@

Equation~\ref{eqn:dtaccel} is supplementary to equation~\ref{eqn:dthydro} in that it
takes into account the possibility of large accelerations causing numerical 
instabilities by violating the Courant condition.  In this equation, $\vec{g}$ is the
gravitational acceleration in each cell on level L.  \blue{ \emph{It should be
  noted that the timestep is computed before the acceleration field is computed,
  and the acceleration field is deleted at the end of each time step.
  So this condition isn't ever actually enforced.  We should either
  fix this problem, or not discuss it.
 -dcollins.} }

>>>>

Algorithmically, the star formation events in Kravtsov's code
are assumed to occur once every global timestep 
$\Delta t_0 \leq 10^7$.  In cells where star formation is
determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
collisionless ``star particle'' is assumed to form, with a 
mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
volume of the mesh cell.  These star particles are
formed instantaneously, with all $m_*$ of gas being
taken out of the cell and immediately deposited into the
star particle.  This particle is then given the velocity
of the gas in the cell, and thereafter treated as a 
collisionless particle.  The \enzo\ implementation
of this algorithm is similar, except that instead of 
forming stars only at the root grid timestep, we allow 
stars to form at the timestep of the highest level of
resolution at any particular point in space.  As can be seen
from the equation for $m_*$ above, this can result in very 
small stellar masses.  To avoid memory and processor time
issues related to having very large numbers of star particles
we impose a threshold mass $M_{*,min}$ such that a star
particle only forms if $m_* \geq M_{*,min}$.  An
appropriate choice of $M_{*,min}$ does not significantly 
change the star overall star formation history of a 
simulation, though it may delay the onset of star formation 
in a given cell relative to a simulation without a particle mass
threshold.

@@

From the point of view of the hydro solver, each grid is solved as an
independent CFD problem, with Dirichlet boundary conditions stored in
the ghost zones.  At the end of each timestep, cells on a given level
fill ghost zones by interpolating from the parent and copying from
neighbors on the same level.  Grids then correct the flux at the interface boundary between
fine and coarse zones, and finally project their real zone data to the
parent grid.

@@

\subsection{Timestepping}\label{sec.timestepping}
\dcc{Made this its own section (not really part of the AMR, it's more
  general than that.)  Added stuff on harmonic average for 3d. See
  note below on acceleration timestep criterion being crap. Moved
  w-cycle to the previous section}
In \enzo, resolution of the equations being solved is adaptive in time 
as well as in space.  The timestep in \enzo\ is satisfied on a level-by-level
 basis by finding the largest timestep such that multiple criteria are
satisfied on each level.  The timestep criteria used by \enzo\ are 
(showing the one-dimensional case for clarity):

@@

Equation~\ref{eqn:dtaccel} is supplementary to equation~\ref{eqn:dthydro} in that it
takes into account the possibility of large accelerations causing numerical 
instabilities by violating the Courant condition.  In this equation, $\vec{g}$ is the
gravitational acceleration in each cell on level L.  \blue{ \emph{It should be
  noted that the timestep is computed before the acceleration field is computed,
  and the acceleration field is deleted at the end of each timestep.
  So this condition isn't ever actually enforced.  We should either
  fix this problem, or not discuss it.
 -dcollins.} }

### 2013-06-06 numerical-radiation-fld.tex
A few more edites to section 2. Also unified "time step" -> "timestep" throughout the whole manuscript.

The system of equations
(\ref{eq:fld_radiation}-\ref{eq:fld_heating}) along with the chemical network 
(Equation~\ref{eq:species_evolution}) is solved
independently of \enzo's hydrodynamics, gravity and dark-matter solvers
(Sections \ref{sec.hydro.ppm}-\ref{sec.ov.nbody}), thereby allowing
the advective portions of Equations (\ref{eq:fld_radiation}) and
(\ref{eq:species_evolution}) to be taken care of by the fluid solvers.
Due to the disparate time scales between radiation transport and chemical
ionization and heating, the remainder of these equations is solved
using an operator-split algorithm.  Within a given time step to evolve
$(E_r^n, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n+1}, {\mathrm
n}_i^{n+1})$, we first evolve equation (\ref{eq:fld_radiation}):
$(E_r^n, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n}, {\mathrm
  n}_i^{n})$.  This uses an implicit Euler time discretization, and a
second-order centered finite-difference spatial discretization,
resulting in a large linear system of equations.  These are solved
using a preconditioned conjugate gradient iteration, where the
preconditioner consists of a geometric multigrid solver.  Both of
these linear solvers are provided by the HYPRE linear solver library
(see \cite{FalgoutYang2002} and \cite{hypre-manual}).

>>>>

The system of equations
(\ref{eq:fld_radiation}-\ref{eq:fld_heating}) along with the chemical network 
(Equation~\ref{eq:species_evolution}) is solved
independently of \enzo's hydrodynamics, gravity and dark-matter solvers
(Sections \ref{sec.hydro.ppm}-\ref{sec.ov.nbody}), thereby allowing
the advective portions of Equations (\ref{eq:fld_radiation}) and
(\ref{eq:species_evolution}) to be taken care of by the fluid solvers.
Due to the disparate time scales between radiation transport and chemical
ionization and heating, the remainder of these equations is solved
using an operator-split algorithm.  Within a given timestep to evolve
$(E_r^n, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n+1}, {\mathrm
n}_i^{n+1})$, we first evolve equation (\ref{eq:fld_radiation}):
$(E_r^n, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n}, {\mathrm
  n}_i^{n})$.  This uses an implicit Euler time discretization, and a
second-order centered finite-difference spatial discretization,
resulting in a large linear system of equations.  These are solved
using a preconditioned conjugate gradient iteration, where the
preconditioner consists of a geometric multigrid solver.  Both of
these linear solvers are provided by the HYPRE linear solver library
(see \cite{FalgoutYang2002} and \cite{hypre-manual}).

### 2013-06-06 overview.tex
A few more edites to section 2. Also unified "time step" -> "timestep" throughout the whole manuscript.

The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
(AMR) capability, which allows it to reach extremely large dynamical
ranges with limited computational resources, opening doors otherwise
closed by finite memory and computational time. Unlike moving mesh
methods \citep{1995ApJS..100..269P,1995ApJS...97..231G} or methods
that subdivide individual cells \citep{Adjerid}, Berger \& Collela's
AMR (also referred to as \emph{structured} AMR) utilizes an adaptive
hierarchy of grid patches at varying levels of resolution.  Each
rectangular grid patch (referred to as a ``grid'') covers some region
of space in its \emph{parent grid} that requires higher resolution,
and can itself become the parent grid to an even more highly resolved
\emph{child grid}.

@@

Given the hierarchy at some time $t$, we advance the solution in the
manner of a W-cycle in a multigrid solver.  First, we determine the
maximum time step allowed for the coarsest grid based on a variety of
accuracy and stability criteria and advance the grid by that time
interval, $\Delta t_0$.  We then move down to the next level and
advance all the grids on that level by a timestep $\Delta t_1$
($\Delta t_1 \leq \Delta t_0$), which is the minimum of all the
allowed timesteps for those grids.  If there are more levels, we
repeat this procedure until the bottom level of the hierarchy has been
reached.  Once there, we continue advancing the grids on the lowest
level until they have ``caught up'' to the coarser parent grids
(i.e. $\sum \Delta t_l = \Delta t_{l-1}$).  This procedure repeats
itself until all grids have been advanced by a total time of $\Delta
t_0$ (see Figure~\ref{fig:wcycle} for a schematic).

@@

Since interesting regions on the grid may move, the hierarchy must
adapt itself. This happens whenever a level has caught up to its
parent level, by entirely rebuilding the grids on that level and at
finer resolutions. Rebuilding is achieved by applying the grid
refinement criteria to the grids on that level, flagging zones that
require extra grids.  These criteria depend on the physical problem
being simulated -- see Section~\ref{sec:refinement_criteria} for more
details.  Once a grid has a set of flagged cells, we apply a
machine-vision based algorithm \citep{Berger91} to find edges and
determine a good placement of subgrids.
%  These subgrids must not overlap one another, must cover all flagged
% cells and their neighbouring cells, and be above a preset efficiency
% threshold, where the efficiency of a cell is defined as the ratio of
% flagged cells to total cells. 
Once these new subgrids have been identified, the solution from the
next coarser grid is interpolated in order to initialize the values on
the new grids.  Finally, any overlap between new subgrids and old ones
is identified, and the prior solution within the regions of overlap is
copied to the new subgrids. This procedure is then repeated on the new
grids and in this way the entire hierarchy (from the original level
examined and continuing on to all finer levels) is rebuilt.

@@

\paragraph{Godunov MUSCL (HD) with Dedner divergence cleaning (MHD)}
This solver was developed to attack problems in magnetic field
amplification during the formation of galaxies \citep{Wang:2009a} and
to understand the role of proto-stellar jets for the theory of star
formation \citep{Wang:2009b}. It combines the standard approach of
Godunov \citep{Godunov1959} for finite volume techniques with the
method of lines as described by \cite{leveque2002finite} and
\cite{toro-1997}. In addition, it implements the hyperbolic divergence
cleaning algorithm of \cite{2002JCoPh.175..645D}. It supports multiple
approximate Riemann solvers and non-ideal equations of
states. Consequently, these suite of solvers can be used for hydro and
magneto-hydrodynamic simulations. This class of solvers, as
well as a version of the PPM hydro solver, has been ported to
nVidia's CUDA framework, allowing \enzo\ to take advantage of modern
graphics hardware \citep{Wang:2010}.

@@

Lastly, we briefly describe the \zeus\ method, a finite-difference
algorithm originally used in the \zeus\ code \citep{Stone92a}. Note
that \enzo\ is entirely independent of the \zeus\ code, and only the
hydrodynamical algorithm of \zeus\ has been implemented in \enzo\ --
the MHD and RHD schemes have not. The \zeus\ method uses a staggered
mesh such that the velocities are face-centered, while the density and
internal energy quantities are cell-centered.  It splits the solution
up into two steps. The first is the so-called source step, in which
the momentum and energy values are updated to reflect the pressure and
gravity forces, including an artificial viscosity required for
stability. The second step, known as the transport step, accounts for
the advection of conserved quantities (mass, momentum and energy)
across the grid.

@@

The current implementation of self-gravity in \enzo\ uses a Fast
Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
coarsest (level 0) grid on each time step.  The advantage of using
this method is that it is fast, accurate, and naturally allows both
periodic and isolated boundary conditions for the gravity, choices
which are very common in astrophysics and cosmology.  On subgrids, we
interpolate the boundary conditions from the parent grid (either the
level 0 grid or some other subgrid). The Poisson equation is then
solved on every timestep using a multigrid technique on one subgrid at
a time. Aside from calculating the gravitational potential arising
from the baryon fields and particles in the simulation, there are also
a number of options for specifying static gravitational fields, and in
fact including self-gravity is optional.

@@

Collisionless matter (e.g., dark matter, stars, etc.) is modeled with
particles that interact with the baryons only via gravity.  These
particles are advanced in a single time step using a drift-kick-drift
algorithm \citep{Hockney88} to provide second-order accuracy even in
the presence of changing timesteps.  Since the particles follow the
collapse of structure, they are not adaptively refined.  Nor are there
duplicate sets of particles for each level; instead, each particle is
associated with the highest refined level available at its position in
the domain and particles are moved between grids as the hierarchy is
rebuilt. Thus, a particle has the same timestep and feels the same
gravitational force as a grid at that refinement level.

Although the particles are fixed in mass once initialized (with the
exception of star particles, which can lose mass in the feedback
process), we can create them with any initial set of masses and
positions.  For example, in many cosmological simulations static
subgrids are included from the beginning in order to improve the
initial spatial and baryonic mass resolution, and these subgrid are
populated with lower mass particles in order to also improve the
collisionless mass resolution. 
%% One particle per initial grid point seems to provide approximately
%% equal sampling between the dark matter and gas. \mqk{This
%% contradicts the conclusions of O'Shea et al. 2005 that an initial
%% grid twice as fine as the mean particle separation is necessary to
%% properly capture the abundance of low mass halos.}

@@

\enzo\ includes the capability of following up to 12 chemical species
using a non-equilibrium solver.  The species can be turned on in sets,
with the simplest model including just H, H$^+$, He, He$^+$, He$^{++}$,
and e$^-$, and more complete models adding first species important for
gas-phase molecular hydrogen formation, H$^-$, H$_2$ and H$_2^+$, and
then HD formation, HD, D, D$^+$. 
%(We note that a reduced version
%of the chemistry, used only with the flux-limited diffusion radiation
%transport, includes only hydrogen species.)
The cooling and heating due to these species is included (see the next
section). The solution of the rate equations is carried out using one
Jacobi iteration of an implicit Euler time discretization to ensure
stability. To ensure accuracy the rate equations are sub-cycled
within one hydrodynamic timestep, subject to the constraint that the
electron and neutral fractions do not change by more than 10\% in one
sub-cycle.

@@

The chemical networks and heating rates described in the previous
sections can be affected by external radiation fields, and the code
includes a number of pre-set radiation backgrounds that are uniform in
space but can vary in time.  These are generally based on the
redshift-dependent rates given in \citet{1996ApJ...461...20H} and
\citet{2012ApJ...746..125H}, but can also include a uniform H$_2$
photo-dissociating background that is either constant in time or
varying as in \citet{WiseAbel05}.

@@

\enzo\ includes a photon-conserving radiative transfer algorithm that
is based on an adaptive ray-tracing method utilizing the HEALPix
pixelization of a sphere \citep{AbelWandelt02}. Photons are integrated
outwards from sources using an adaptive time-stepping scheme that
preserves accuracy in ionization fronts even in the optically-thin
limit. This has been coupled to the chemistry and cooling network to
provide ionization and heating rates on a cell-by-cell basis. The
method is described in detail (including numerous code tests) in
\citet{Wise11_Moray}.

@@

Heat conduction, both isotropic and anisotropic, can be included using
a sub-cycled, operator-split method.  The heat fluxes are computed
with simple second-order accurate finite differences, and stability is
ensured by restricting the time step and using flux-limiters where
appropriate.

@@

All grids on a given level are advanced with the same timestep,
determined by calculating, for each cell and for each physical process
(except for the chemistry step, which is sub-cycled) separately, the
minimum $\Delta t$ allowed for stability. The level is advanced with a
timestep equal to the minimum over all of these $\Delta t$.

>>>>

The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
(AMR) capability, which allows it to reach extremely large dynamical
ranges with limited computational resources, opening doors otherwise
closed by finite memory and computational time. Unlike moving mesh
methods \citep{1995ApJS..100..269P,1995ApJS...97..231G} or methods
that subdivide individual cells \citep{Adjerid}, Berger \& Collela's
AMR (also referred to as \emph{structured} AMR) utilizes an adaptive
hierarchy of grid patches at varying levels of resolution.  Each
rectangular grid patch (referred to as a ``grid'') covers some region
of space in its \emph{parent grid} requiring higher resolution,
and can itself become the parent grid to an even more highly resolved
\emph{child grid}.

@@

Given the hierarchy at some time $t$, we advance the solution in the
manner of a W-cycle in a multigrid solver.  First, we determine the
maximum timestep allowed for the coarsest grid based on a variety of
accuracy and stability criteria and advance the grid by that time
interval, $\Delta t_0$.  We then move down to the next level and
advance all the grids on that level by a timestep $\Delta t_1 \leq
\Delta t_0$, which is the minimum of all the allowed timesteps for
those grids.  If there are more levels, we repeat this procedure until
the bottom level of the hierarchy has been reached.  Once there, we
continue advancing the grids on the lowest level until they have
``caught up'' to the coarser parent grids (i.e. $\sum \Delta t_l =
\Delta t_{l-1}$).  This procedure repeats itself until all grids have
been advanced by a total time of $\Delta t_0$ (see
Figure~\ref{fig:wcycle} for a schematic).

@@

Since interesting regions on the grid may move, the hierarchy must
adapt itself. This happens whenever a level has caught up to its
parent level, by entirely rebuilding the grids on that level and at
finer resolutions. Rebuilding is achieved by applying the grid
refinement criteria to the grids on that level, flagging zones that
require extra grids.  These criteria depend on the physical problem
being simulated; see Section~\ref{sec:refinement_criteria} for more
details.  Once a grid has a set of flagged cells, we apply a
machine-vision based algorithm \citep{Berger91} to find edges and
determine a good placement of subgrids.
%  These subgrids must not overlap one another, must cover all flagged
% cells and their neighbouring cells, and be above a preset efficiency
% threshold, where the efficiency of a cell is defined as the ratio of
% flagged cells to total cells. 
Once these new subgrids have been identified, the solution from the
next coarser grid is interpolated in order to initialize the values on
the new grids.  Finally, any overlap between new subgrids and old ones
is identified, and the prior solution within the regions of overlap is
copied to the new subgrids. This procedure is then repeated on the new
grids and in this way the entire hierarchy (from the original level
examined and continuing on to all finer levels) is rebuilt.

@@

\paragraph{Godunov MUSCL (HD) with Dedner divergence cleaning (MHD)}
This solver was developed to attack problems in magnetic field
amplification during the formation of galaxies \citep{Wang:2009a} and
to understand the role of proto-stellar jets for the theory of star
formation \citep{Wang:2009b}. It combines the standard approach of
Godunov \citep{Godunov1959} for finite volume techniques with the
method of lines as described by \cite{leveque2002finite} and
\cite{toro-1997}. In addition, it implements the hyperbolic divergence
cleaning algorithm of \cite{2002JCoPh.175..645D}. It supports multiple
approximate Riemann solvers and non-ideal equations of
states. Consequently, this suite of solvers can be used for hydro and
magneto-hydrodynamic simulations. This class of solvers, as
well as a version of the PPM hydro solver, has been ported to
nVidia's CUDA framework, allowing \enzo\ to take advantage of modern
graphics hardware \citep{Wang:2010}.

@@

Lastly, we briefly describe the \zeus\ method, a finite-difference
algorithm originally used in the \zeus\ code \citep{Stone92a}. Note
that \enzo\ is entirely independent of the \zeus\ code, and only the
hydrodynamical algorithm of \zeus\ has been implemented in \enzo; the
MHD and radiation-hydrodynamics schemes have not. The \zeus\ method
uses a staggered mesh such that velocities are face-centered, while
density and internal energy are cell-centered.  It splits the solution
up into two steps. The first is the so-called source step, in which
the momentum and energy values are updated to reflect the pressure and
gravity forces, including an artificial viscosity required for
stability. The second step, known as the transport step, accounts for
the advection of conserved quantities (mass, momentum, and energy)
across the grid.

@@

The current implementation of self-gravity in \enzo\ uses a Fast
Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
root grid on each timestep.  The advantage of using this method is
that it is fast, accurate, and naturally allows both periodic and
isolated boundary conditions for the gravity, choices which are very
common in astrophysics and cosmology.  On subgrids, we interpolate the
boundary conditions from the parent grid (either the root grid or some
other subgrid). The Poisson equation is then solved on every timestep
using a multigrid technique on one subgrid at a time. Aside from
self-consistently calculating the gravitational potential arising from
the baryon fields and particles in the simulation, there are also a
number of options for specifying static gravitational fields, and in
fact including self-gravity is optional.

@@

Collisionless matter (e.g. dark matter, stars, etc.) is modeled with
particles that interact with the baryons only via gravity.  These
particles are advanced through a single timestep using a drift-kick-drift
algorithm \citep{Hockney88} to provide second-order accuracy even in
the presence of varying timesteps.  Since the particles follow the
collapse of structure, they are not adaptively refined.  Nor are there
duplicate sets of particles for each level; instead, each particle is
associated with the highest refined level available at its position in
the domain and particles are moved between grids as the hierarchy is
rebuilt. Thus a particle has the same timestep and feels the same
gravitational force as a grid at that refinement level.

Although the particles are fixed in mass once initialized (with the
exception of star particles, which can lose mass in the feedback
process), we can create them with any initial set of masses and
positions.  For example, in many cosmological simulations static
subgrids are included from the beginning in order to improve the
initial spatial and baryonic mass resolution, and these subgrid are
populated with lower mass particles to correspondingly improve the
collisionless mass resolution.
%% One particle per initial grid point seems to provide approximately
%% equal sampling between the dark matter and gas. \mqk{This
%% contradicts the conclusions of O'Shea et al. 2005 that an initial
%% grid twice as fine as the mean particle separation is necessary to
%% properly capture the abundance of low mass halos.}

@@

\enzo\ includes the capability of following up to 12 particle species
using a non-equilibrium chemical network.  The species can be turned
on in sets, with the simplest model including just atomic hydrogen and
helium (H, H$^+$, He, He$^+$, He$^{++}$, e$^-$), and more complete
models adding first species important for gas-phase molecular hydrogen
formation (H$^-$, H$_2$ and H$_2^+$), and then HD formation (HD, D,
D$^+$).
%(We note that a reduced version
%of the chemistry, used only with the flux-limited diffusion radiation
%transport, includes only hydrogen species.)
The cooling and heating due to these species is included (see the next
section). The solution of the rate equations is carried out using one
Jacobi iteration of an implicit Euler time discretization to ensure
stability. To ensure accuracy the rate equations are sub-cycled
within one hydrodynamic timestep, subject to the constraint that the
electron and neutral fractions do not change by more than 10\% in one
sub-cycle.

@@

The chemical networks and heating rates described in the previous
sections can be affected by external radiation fields, and the code
includes a number of pre-calculated meta-galactic UV radiation
backgrounds that are uniform in space but can vary in time.  These are
generally based on the redshift-dependent rates given in
\citet{1996ApJ...461...20H} and \citet{2012ApJ...746..125H}, but can
also include a uniform H$_2$ photo-dissociating background that is
either constant in time or varying as in \citet{WiseAbel05}.

@@

\enzo\ includes a photon-conserving radiative transfer algorithm that
is based on an adaptive ray-tracing method utilizing the HEALPix
pixelization of a sphere \citep{AbelWandelt02}. Photons are integrated
outwards from sources using an adaptive timestepping scheme that
preserves accuracy in ionization fronts even in the optically-thin
limit. This has been coupled to the chemistry and cooling network to
provide ionization and heating rates on a cell-by-cell basis. The
method is described in detail (including numerous code tests) in
\citet{Wise11_Moray}.

@@

Heat conduction, both isotropic and anisotropic, can be included using
a sub-cycled, operator-split method.  The heat fluxes are computed
with simple second-order accurate finite differences, and stability is
ensured by restricting the timestep and using flux-limiters where
appropriate.

@@

All grids on a given level are advanced with the same timestep,
determined by calculating, for each cell and for each physical process
(except for chemistry and heat conduction, which are sub-cycled)
separately, the largest $\Delta t$ allowed for stability. The level is
advanced with a timestep equal to the minimum over all of these
$\Delta t$.
